Runaway and Hypervelocity Stars from Compact Object Encounters in Globular Clusters

### 2023-02-08 src/tex/ms.tex
Fixed some references

Alternate acceleration mechanisms capable of producing high-velocity stars include the binary supernova and dynamical ejection scenarios (BSS \citep{1961BAN....15..265B} and DES \citep{1967BOTT....4...86P}, respectively).
In the BSS, the more massive primary of a stellar binary undergoes a supernova, which subsequently accelerates the companion; this scenario has been predicted to accelerate stars to speeds of a few hundred \kms \citep{2019A&A...624A..66R, 2022MNRAS.tmp.3236I}, with exceptionally light companions potentially receiving velocities in excess of 1000 \kms \citep{2015MNRAS.448L...6T}.
The DES concerns strong gravitational interactions among 3-4 stellar objects, and has been associated with higher speed limits: \citet{1991AJ....101..562L} studied these encounters with numerical methods, and found that the upper limit for products of these encounters was approximately the escape speed from the surface of the most massive star involved.
For a sun-like star, this escape speed is $\sim$620 \kms, while for a 60 $M_\odot$ late main-sequence star it is $\sim$1400 \kms.
\citet{2023arXiv230111444O} recently found that for OB runaways in the Small Magellanic Cloud, the DES is expected to occur 2-3 times more frequently that the BSS.

Globular clusters (GCs) are obvious candidate matrices for both of these events due to their high stellar densities, but the DES might be exceptionally amplified due to the presence of BH subsystems in the centers of GCs.
Multi-band observations \citep{Maccarone2007a, Barnard2011, Chomiuk2013, Miller-Jones2015a} employing a variety of techniques \citep{Strader2012, Bahramian2016, Giesers2018} have been successful in finding tens of BH candidates in GCs, which are in themselves consistent with the presence of hundreds of BHs per cluster \citep{Giesers2019}.
BHs play important roles in the macroscopic evolution of the GC \citep[e.g.][]{2020IAUS..351..357K} by dominating gravitational interactions in GC cores; this naturally places BHs in a privileged position when considering the DES in GCs, particularly when considering the maximum velocities attainable by this mechanism.

Several recent and ongoing studies seek investigate these phenomena and the broader question of extra-tidal stars from GCs, in part encouraged by the continual improvement of astrometric measurements through efforts such as \textit{Gaia} \citet{2022A&A...667A.148G}.
\citet{2023MNRAS.518.4249G} developed and used a particle spray code to model ejections from GC cores and demonstrate that kinematic cuts can be too aggressive when searching for ejecta, preferring the use of chemical abundances alone.
\citet{2023arXiv230105166F} presented a comprehensive catalog of extra-tidal features of MWGCs, produced by simulating tidal stripping of known MWGCs in the context of the MW potential.
Of particular proximity to our work, \citet{2022arXiv221116523W} used the same GC model catalog we employ to holistically examine stellar ejections from GCs and identify key mechanisms.
The latter two of of these works are explicitly presented as the first in a series of papers seeking to compose a more complete picture of their respective objects.

This work investigates the capacity of GCs to produce HVSs through the DES.
We focus on encounter between binary and single objects, as they are the most abundant kind of strong encounter, and specifically those that involve compact objects (COs, meaning BHs, a white dwarfs (WDs), and neutron stars (NSs)).
In \S\ref{sec:methods}, we describe our method of generating binary-single compact object (BSCO) ejecta populations from realistic simulations of GCs.
In \S\ref{sec:gcdyns}, we examine these populations as functions of GC parameters, and the relationship they have to GC evolution.
In \S\ref{sec:est_MW-like}, we combine our synthetic populations with observational MWGC catalogs to produce a MW-like population of ejecta, along with predicting rates and phase space distributions for these objects.
We conclude that GCs are possible generators of HVSs in all velocity regimes thus observed, beyond the previously established limit for star-only encounters, albeit the rate at which these objects are produced from GCs is significantly lower than that for the galactic center.

@@

\subsection{GC core collapses} \label{subsec:cc}

Core collapse is a well-documented feature of GC evolution (e.g. \citealt{1968MNRAS.138..495L}, \citealt{2001A&A...375..711F}, \citealt{2006MNRAS.368..121F}, \citealt{2008gady.book.....B}, \citealt{2020IAUS..351..357K}), and as will be seen is a significant factor in the ejection of stars from these systems.
In general, a core collapse occurs when gravitational potential energy is rapidly liberated from the central region of a GC through dynamical encounters, causing the core to contract while energizing and potentially expanding the surrounding envelope of stars.
Binaries play an important role in these processes, as their formation and hardening through strong encounters accelerate the other interacting stars, which subsequently travel to the more extended regions of the GC and potentially escape it entirely.
The first core collapse of a GC tends to be propagated by these dynamics, causing the GC to puff up over the following Gyrs as the central binaries continue to shrink and the stellar density increases.

Dynamical friction-induced mass segregation \citep[e.g.][]{2008gady.book.....B} causes the core to be populated by the most massive objects in the GC, which also are those that first evolve into BHs.
As the core contracts, successively stronger encounters eject all but a small handful of the initially 10s-100s of BHs; this reduced subsystem is unable to support the GC to the same extent as its predecessor, and so the interior of the GC contracts until a sufficiently strong replacement is composed.
After \citet{2020IAUS..351..357K}, we employ this characteristic of BH ejection as a metric of whether a GC model has undergone this second kind of core collapse: once there are less than 10 BHs remaining in the model, we label it as core collapsed from that time onward.

@@

The 148 \CMC\ models are designated by cluster initial size $N$ (number of stellar objects), initial virial radius $r_{\rm vir}$ (parsecs), distance from the galactic center $r_{\rm gc}$ (kiloparsecs, used in \CMC\ to calculate tidal effects from the galactic potential), and metallicity $Z$ (used to prescribe star evolution), with values chosen to span much of the MWGC parameter space; see \citet{2020IAUS..351..357K} for more details.
To get a clear idea of how cluster evolution is linked to BSCO ejections, we examine four sample models; our base \CMCcat\ model has initial conditions
\begin{itemize}
    \item $N = 8 \times 10^5$
    \item $r_{\rm vir} = 0.5~{\rm pc}$
    \item $r_{\rm gc} = 8~{\rm kpc}$
    \item $Z = 0.01Z_\odot$,
\end{itemize}
which, except for the compact virial radius, are near the typical values for the MWGC population (the small virial radius model is chosen for its shorter relaxation time, promising a greater dynamical age for a given physical time).
Each of the other three sample models vary one of these parameters (either $N \to 4 \times 10^5$, $r_{\rm vir} \to 2~{\rm pc}$, or $Z \to Z_\odot$) to promote understanding of the effects of each.

Figure \ref{fig:cmc_single_clusters} shows the ejection velocity $v_{\rm out}$ for every MS star ejected via.~a BSCO encounter in these models, plotted at the time of encounter from initialization $t$.
What is immediately apparent from these single cluster models is a strong correlation between BSCO encounters and the evolution of the cluster core: whenever a cluster undergoes a core collapse, the heightened densities lead to predictably high encounter rates.

@@

An interplay between stellar evolution and cluster dynamics is revealed in the kinds of BSCO encounters that occur at certain times.
The most massive stars in a cluster are both the first to form binaries and the first to evolve into COs; hence, encounters between a mixed binary (composed of a MS star and a CO) and a single MS star tend to dominate the first generations of BSCO encounters.
This dominance usually continues until the first core collapse after CO formation, where the heightened densities facilitate a rise in encounters involving multiple COs.
There is a stark difference in the ejection velocities produced by these encounters in comparison to the previous mixed binary-single MS star encounters, and the fastest ejections over the cluster's lifetime are produced in the first tens of Myr after the associated ``first" core collapse.
This regime lasts for a few to tens of Gyr, during which the COs in the core (predominantly BHs) preferentially form CO-CO binaries, as evidenced in the drop of mixed binary-single CO encounters during this phase.

For cluster models that evolve to BH-depletion ``second" core collapse (in our sample, all but the large model), the corresponding increase in BSCO encounters is dominated by stellar binary-single CO encounters.
We also find that the majority of these encounters do not involve BHs, but rather WDs (or in some cases a NSs).
These behaviors are consistent with mass segregation, where it is only after the higher-mass BHs are ejected that these lower-mass COs can migrate into the core.

@@

The $r_{\rm vir} = 2 {\rm pc}$ model evolves to ``first" core collapse later than the previous two, and noticeably lacks the pre-core collapse abundance of mixed binary-single star encounters seen above.
When ``first" core collapse does occur, the distribution of encounters is similar to the respective distributions for the base model.

The difference between the solar metallicity model and the others is striking.
This model does not evolve to ``first" core collapse until $\sim$200 Myr (by far the latest out of the four), but still reaches ``second" core collapse near 7 Gyr.
Furthermore, the lower stellar masses make the first core collapse much less severe, delaying the maximum ejecta production rate until the second core collapse; this rate increases roughly monotonically throughout the model's evolution.

The $r_{\rm vir} = 2 pc$ model evolves slower, and so it does not reach the same core collapse phase as the previous two.
The lower densities prevent the initial dominance of mixed binary-single star encounters, as the frequency of BSCO encounters is relatively low until the initial core collapse.
On the other hand, this first core collapse is not associated with an abundance of BSCO encounters in the high-metallicity model, which sees a roughly monotonic rise in BSCO encounters throughout its evolution.

The dominance of the core in overall BSCO encounter production is even more clear when considering the radial localization of the encounters.
\CMC\ does not record the radius where each strong encounter occurred, but it does record the local escape velocity.
For comparison, the escape velocity from the center of the model can be computed from the central and tidal boundary potentials, which are recorded throughout the evolution of the cluster.\footnote{These potentials can be found in the standard \CMC\ output file \texttt{initial.esc.dat}, which records data pertaining to stars that become unbound to the model for any reason.  The escape rate is high enough that the time domain is well-resolved by these data.}
Figure \ref{fig:cmc_single_clusters_vesc} plots these data for the same sample clusters, where the distribution of encounter escape velocities is strikingly correlated to the central escape velocity, indicating the degree to which the core dominates these dynamics.

@@

The most noticeable distinction among models of different metallicities is the number of ejections, which decreases with increasing metallicity.
This is understandable, as in practice the lower metallicity models have greater BH populations than the higher metallicity models; quantitatively, the 1\% solar models have about $1.2 \times$ as many BHs at the time of ``first" core collapse as the solar models, and maximum BH masses about $2 \times$ those of the same.
This weakens BSCO dynamics before and during ``first" core collapse, as was first made visible in Figure \ref{fig:cmc_single_clusters}.
Because this effect takes place in the earlier stages of model evolution - namely, when mass segregation has had less time to separate lighter objects from the strong dynamics of heavier objects - this has the most significant effect of reducing the number of ejecta with $m < M_\odot$.
Notably, the different metallicity models have quite similar ejecta mass distributions after ``second" core collapse, as the effect of metallicity on the masses of the remaining WDs and NSs is much less pronounced than on the now-ejected BHs.

@@

Having constrained the snapshots by $r_{\rm gc}$ and metallicity, we then choose one of an appropriate size and evolutionary state by comparing the snapshots with the model in normalized $\log M$ - $r_c / r_h$ space, where $M$ is the mass of the cluster in $M_\odot$, $r_c$ is the Spitzer core radius \citep{1987degc.book.....S}
\begin{equation}
    r_c = \sqrt{\frac{3 \sigma_c^2}{4 \pi \rho_c}},
\end{equation}
(where $\sigma_c$ is the central velocity dispersion and $\rho_c$ is the central density) and $r_h$ is the 3D half-mass radius of the snapshot (the latter two parameters are used because both are immediately accessible in the \citet{2018MNRAS.478.1520B} catalog and standard \CMC\ output).
The models and GCs are plotted in the non-normalized space of these parameters in the bottom plot of Figure \ref{fig:cmcs-mwgcs_scatter}.
The transformation to the normalized space is simply
\begin{equation}
    x_{\rm norm} = \frac{x - \bar{x}_{\rm \CMC}}{\sigma_{\rm x, \CMC}},
\end{equation}
where $\bar{x}_{\rm CMC}$ and $\sigma_{\rm x,CMC}$ are respectively the mean and standard deviation of the parameter $x$ over the complete set of \CMC\ snapshots.
The snapshot from the $r_{\rm gc}$ - [Fe/H] subset that is closest to the MWGC in the normalized $\log M$ - $r_c / r_h$ space is chosen to represent the GC; as an example, 47 Tuc ($r_{\rm gc} = 7.52$ kpc, $Z = 0.0019$; $\log M = 5.95$, $r_c / r_h,m = 0.125$) is represented by a snapshot from model \texttt{N1.6e6\_rv2\_rg8\_Z0.002}, which at the 10.8 Gyr time of the snapshot has $\log M = 5.91$ and $r_c / r_h = 0.126$.

@@

Having chosen representative snapshots for all MWGCs, we now use our synthetic ejecta populations to compose a MW-like population of BSCO ejecta from GCs.
All orbit integrations in the galactic frame are done with \texttt{galpy}\footnote{http://github.com/jobovy/galpy} \citep{2015ApJS..216...29B}.

Setting the snapshot to the present-day galactic orbital parameters from \citet{2018MNRAS.478.1520B}, we back-integrate the orbit of each GC in the \texttt{galpy}\texttt{MWPotential2014} potential to the initialization time of the respective model.
We then locate each ejected object at the appropriate place in the GC orbit, using the time of BSCO encounter (naturally ignoring any ejections that occurred after the time of the chosen snapshot).
A random ejection direction is chosen in the GC rest frame for each object, and the velocities are then transformed to the galactic rest frame.
Finally, the orbits of all ejecta are then integrated to the present day in the same galactic potential.

@@

\begin{figure*}
    \script{gc_orbit_ejections.py}
    \gridline{
        \fig{figures/gc_orbit_ejections_NGC_104.pdf}{0.49\textwidth}{}
        \fig{figures/gc_orbit_ejections_NGC_5139.pdf}{0.49\textwidth}{}
    }
    \gridline{
        \fig{figures/gc_orbit_ejections_NGC_6205.pdf}{0.49\textwidth}{}
        \fig{figures/gc_orbit_ejections_NGC_7089.pdf}{0.49\textwidth}{}
    }
    \caption{
        Plots showing the back-integrated orbits for some sample MWGCs (gray curves), and the points in the orbit where an object is ejected from the GC (scatter points).
        The color scale communicates the mass of the ejected star.
        The concentration of ejections in the first few orbits is clear, and an increased density of ejections when clusters are farther away from the galactic center is visible as well.
        Viewing the figure electronically makes it easier to find the few high-mass ejecta amid the abundance of lower mass objects.
    }
    \label{fig:gc_orbit_ejections}
\end{figure*}

@@

From a galactocentric perspective, the population of runaways is fairly isotropic in position, as can be seen in Figure \ref{fig:rgc-z}.
The distributions of distance from the galactic origin, and the local distributions of galactocentric velocity are similar when distributing over radius $r_{\rm gc}$ versus distance from the galactic plane $Z_{\rm gc}$.
Most runaways end up at a distance on the order of 10 pc from the galactic center, and with a velocity on the order of 100-300~\kms.
Few runaways make it past the $\sim$100 pc mark, but those that do naturally retain the highest velocities, reaching upwards of 1000~\kms.

\begin{figure*}
    \script{rgc-z.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/rgc-z_S.pdf}
        \caption{
            Histograms and velocity quantiles for the synthetic ejecta.
            The left (right) plots show the profile over radial distance from the galactic center $r_{\rm gc}$ (distance from the galactic plane $Z$).
            The quantiles in the lower plots are calculated from the present-day velocities of our population.
        }
        \label{fig:rgc-z}
    \end{centering}
\end{figure*}

One result of note is that of the heliocentric radial velocity $v_{\rm rf}$ distribution of the synthetic population.
\citet{2021arXiv211213864G} in part studied the galactic center origin of HVSs, and found that there is an apparent tension between the observed and predicted runaway populations; specifically, the predicted number of stars with velocities $\gtrsim 700~{\rm km~s^{-1}}$ was much higher than the observed rate from the HVS sample of \citet{2018ApJ...866...39B}.
The same work found that a decreasing star formation rate over last millions of years would reduce the recent (and thus more likely to be observed) HVS production rate enough to overcome the discrepancy.
GCs naturally follow this pattern (c.f. \S\ref{sec:gcdyns}), as the BSCO runaways studied here are produced at much higher rates when the models are young, and few HVSs are produced in the later stages of the cluster.
The resulting $v_{\rm rf}$ distribution is much closer to the observed distribution than that for the galactic center origin case, albeit skewed towards slightly smaller velocities (Figure \ref{fig:cmc_orbits_cdf}).
This latter difference grows when comparing to the HVS sample of \citet{2018ApJ...866..121H}, who focused on metal-poor stars, which are more akin to the objects that populate GCs.

\begin{figure}
    \script{cmc_orbits_cdf.py}
    \begin{centering}
        \includegraphics[width=0.99\linewidth]{figures/cmc_orbits_cdf_S.pdf}
        \caption{
            Velocity distributions for our sythetic and some previous observational HVS catalogs.
            $v_{\rm rf}$ is the heliocentric radial velocity of the star as measured in the galactic rest frame.
        }
        \label{fig:cmc_orbits_cdf}
    \end{centering}
\end{figure}

Figure \ref{fig:mwej_rates} shows the rate of HVS/runaway production for our synthetic population, along with some base values from the literature.
In both cases, the GC ejection rate is smaller than the observed rate, especially at the present day.
However, runaways from GCs could make up an appreciable fraction of runaways in the early universe.
It is also of note that GCs are able to produce any HVSs at all, at velocities previously associated entirely with the galactic center.

>>>>

Alternate acceleration mechanisms capable of producing high-velocity stars include the binary supernova and dynamical ejection scenarios (BSS \citep{1961BAN....15..265B} and DES \citep{1967BOTT....4...86P}, respectively).
In the BSS, the more massive primary of a stellar binary undergoes a supernova, which subsequently accelerates the companion; this scenario has been predicted to accelerate stars to speeds of a few hundred \kms \citep{2019A&A...624A..66R, 2022MNRAS.tmp.3236I}, with exceptionally light companions potentially receiving velocities in excess of 1000 \kms \citep{2015MNRAS.448L...6T}.
The DES concerns strong gravitational interactions among 3-4 stellar objects, and has been associated with higher speed limits: \citet{1991AJ....101..562L} studied these encounters with numerical methods, and found that the upper limit for products of these encounters was approximately the escape speed from the surface of the most massive star involved.
For a sun-like star, this escape speed is $\sim$620 \kms, while for a 60 $M_\odot$ late main-sequence star it is $\sim$1400 \kms.
\citet{2020ApJ...903...43D} found that for OB runaways in the Small Magellanic Cloud, the DES is expected to occur 2-3 times more frequently that the BSS.

Globular clusters (GCs) are obvious candidate matrices for both of these events due to their high stellar densities, but the DES might be exceptionally amplified due to the presence of BH subsystems in the centers of GCs.
Multi-band observations \citep{Maccarone2007a, Barnard2011, Chomiuk2013, Miller-Jones2015a} employing a variety of techniques \citep{Strader2012, Bahramian2016, Giesers2018} have been successful in finding tens of BH candidates in GCs, which are in themselves consistent with the presence of hundreds of BHs per cluster \citep{2018ApJ...855L..15K, Giesers2019}.
BHs play important roles in the macroscopic evolution of the GC \citep[e.g.][]{2020IAUS..351..357K} by dominating gravitational interactions in GC cores; this naturally places BHs in a privileged position when considering the DES in GCs, particularly when considering the maximum velocities attainable by this mechanism.

Several recent and ongoing studies have investigated these phenomena and the broader question of extra-tidal stars from GCs, in part encouraged by the continual improvement of astrometric measurements through efforts such as \textit{Gaia} \citep{2022arXiv220800211G}.
\citet{2023MNRAS.518.4249G} developed and used a particle spray code to model ejections from GC cores and demonstrate that kinematic cuts can be too aggressive when searching for ejecta, preferring the use of chemical abundances alone.
\citet{2023arXiv230105166F} presented a comprehensive catalog of extra-tidal features of MWGCs, produced by simulating tidal stripping of known MWGCs in the context of the MW potential.
Of particular proximity to our work, \citet{2022arXiv221116523W} used the same GC model catalog we employ to holistically examine stellar ejections from GCs and identify key mechanisms.
The latter two of of these works are explicitly presented as the first in a series of papers seeking to compose a more complete picture of their respective objects.

This work investigates the capacity of GCs to produce HVSs through the DES.
We focus on encounters between binary and single objects, as they are the most abundant kind of strong encounter, and specifically those that involve compact objects (COs, meaning BHs, a white dwarfs (WDs), and neutron stars (NSs)).
In \S\ref{sec:methods}, we describe our method of generating binary-single compact object (BSCO) ejecta populations from realistic simulations of GCs.
In \S\ref{sec:gcdyns}, we examine these populations as functions of GC parameters, and the relationship they have to GC evolution.
In \S\ref{sec:est_MW-like}, we combine our synthetic populations with observational MWGC catalogs to produce a MW-like population of ejecta, along with predicting rates and phase space distributions for these objects.
We conclude that GCs are possible generators of HVSs in all velocity regimes thus observed, beyond the previously established limit for star-only encounters, albeit the rate at which these objects are produced from GCs is significantly lower than that for the galactic center.

@@

\subsection{Core collapse in GCs} \label{subsec:cc}

Core collapses are well-documented features of GC evolution \citep[e.g.][]{1968MNRAS.138..495L, 2001A&A...375..711F, 2006MNRAS.368..121F, 2008gady.book.....B, 2020IAUS..351..357K}, and the heightened densities during such phenomena makes them relevant to the present study.
This process occurs due to the energy transport by two-body relaxation from the central regions of the cluster outwards, causing a core to develop and contract while the outer regions expand \citep{2003gmbp.book.....H}.
The presence of massive objects (e.g. BHs) can significantly accelerate this process, as dynamical friction slows the most massive objects in the system, causing them to segregate into the core on a much more rapid timescale \citep{2008gady.book.....B}.
Following \cite{2013MNRAS.432.2779B}, we refer to this initial BH-lead collapse as the ``first’’ core collapse.
While this process can be effected substantially by the metallicity of the star cluster, with more metal-rich clusters experiencing greater mass loss due to stellar winds and producing lower-mass BHs with correspondingly longer dynamical friction timescales (fewer of which are retained due to natal kicks) \citep{2022arXiv220316547R}, core collapse typically occurs within a few 100 Myr for most GCs.

% From the beginning of a GC's lifetime, dynamical friction slows the most massive objects in the system, while simultaneously accelerating the less massive objects \citep[e.g.][]{2008gady.book.....B}.
% Globally, this leads to the expansion of the outer layers of the cluster with a contemporary contraction of the core \citep{2013MNRAS.429.2881C}, usually within 100 Myr of the birth of the system.

%This first core collapse is delayed and weakened in systems with higher metallicities: the increased mass loss due to stronger stellar winds directly leads to smaller objects that experience weaker dynamical friction, and indirectly the number of ``massive" objects is reduced as any that to evolve into BHs are more easily removed from the GC by natal kicks \citep{2008MNRAS.386...65M, 2022arXiv220316547R}.

First core collapse is stopped by the formation and hardening of binaries in the core (which at this point is largely composed of BHs), and the liberated energy supports the universal expansion of the GC for a time on the order of Gyr \citep{2012MNRAS.420..309B}.
The BH population in the core is gradually depleted as strong encounters accelerate their participants beyond the local escape speed, until no more than a handful of BHs remain in the cluster \citep{2013MNRAS.432.2779B, 2020IAUS..351..357K}.
The loss of the dynamical heat source causes the core to contract again, as the system adjusts to being supported by WDs, NSs and stars in the same mass regime.
Unlike the first collapse, this second, permanent core collapse is observable in that the surface brightness profile of the GC becomes cuspy to the limit of the cluster center (and is what observers would classically call a ``core-collapsed cluster'').
We follow \citet{2018ApJ...855L..15K} in pairing these demographic and observable features, marking the time of second core collapse in our models as the same time the number of BHs drops below 10.

@@

The 148 \CMC\ models are designated by cluster initial size $N$ (number of stellar objects), initial virial radius $r_{\rm vir}$ (parsecs), distance from the galactic center $r_{\rm gc}$ (kiloparsecs, used in \CMC\ to calculate tidal effects from the galactic potential), and metallicity $Z$ (used to prescribe star evolution), with values chosen to span much of the MWGC parameter space; see \citet{2020IAUS..351..357K} for more details.
To get a clear idea of how cluster evolution is linked to BSCO ejections, we examine four sample models; our base \CMCcat\ model has initial conditions
\begin{itemize}
    \item $N = 8 \times 10^5$
    \item $r_{\rm vir} = 0.5~{\rm pc}$
    \item $r_{\rm gc} = 8~{\rm kpc}$
    \item $Z = 0.01Z_\odot$.
\end{itemize}
Each of the other three sample models vary one of these parameters (either $N \to 4 \times 10^5$, $r_{\rm vir} \to 2~{\rm pc}$, or $Z \to Z_\odot$) to promote understanding of the effects of each.
The base $N$, $r_{\rm gc}$, and $Z$ were chosen as they correspond to typical values for MWGCs; the notably small $r_{\rm vir}$ is chosen for the samples because the dynamic range of cluster evolution and stellar ejections is greatest among \CMC\ models of this size, making it easier to identify the same features present in larger models.

Figure \ref{fig:cmc_single_clusters} shows the ejection velocity $v_{\rm out}$ for every MS star ejected through a BSCO encounter in these models, plotted at the time of encounter from initialization $t$.
What is immediately apparent from these single cluster models is a strong correlation between BSCO encounters and the evolution of the cluster core: whenever a cluster undergoes a core collapse, the heightened densities lead to predictably high encounter rates.

@@

An interplay between stellar evolution and cluster dynamics is revealed in the kinds of BSCO encounters that occur at certain times.
The most massive stars in a cluster are both the first to form binaries and the first to evolve into COs; hence, encounters between a mixed binary (composed of a MS star and a CO) and a single MS star tend to dominate the first generations of BSCO encounters.
This dominance usually continues until the first core collapse after CO formation, where the heightened densities facilitate a rise in encounters involving multiple COs.
There is a stark difference in the ejection velocities produced by these encounters in comparison to the previous mixed binary-single MS star encounters, and the fastest ejections over the cluster's lifetime are produced in the first tens of Myr after the associated first core collapse.
This regime lasts for a few to tens of Gyr, during which the COs in the core (predominantly BHs) preferentially form CO-CO binaries, as evidenced in the drop of mixed binary-single CO encounters during this phase.

For cluster models that evolve to BH-depletion second core collapse (in our sample, all but the $r_{\rm vir} = 2~{\rm pc}$ model), the corresponding increase in BSCO encounters is dominated by stellar binary-single CO encounters.
We also find that the majority of these encounters do not involve BHs, but rather WDs (or in some cases a NSs).
These behaviors are consistent with mass segregation, where it is only after the higher-mass BHs are ejected that these lower-mass COs can migrate into the core.

@@

The $r_{\rm vir} = 2~{\rm pc}$ model evolves to first core collapse later than the previous two, and noticeably lacks the pre-core collapse abundance of mixed binary-single star encounters seen above.
When first core collapse does occur, the distribution of encounters is similar to the respective distributions for the base model.

The difference between the solar metallicity model and the others is striking.
This model does not evolve to first core collapse until $\sim$200 Myr (by far the latest out of the four), and leads to typical densities no more than 10\% those of the other models; this occurs due to the lower BH masses and abundances associated with the greater mass loss (c.f. \S\ref{subsec:cc} above, and \S5.2 in \citet{2022arXiv220316547R}).
Also unlike the other models, here the number of ejections does not peak until the second, BH-depletion core collapse near 7 Gyr.

The dominance of the core in overall BSCO encounter production is even more clear when considering the radial localization of the encounters.
\CMC\ does not record the radius where each strong encounter occurred, but it does record the local escape velocity.
For comparison, the escape velocity from the center of the model can be computed from the central and tidal boundary potentials, which are recorded throughout the evolution of the cluster.
Figure \ref{fig:cmc_single_clusters_vesc} plots these data for the same sample clusters, where the distribution of encounter escape velocities is strikingly correlated to the central escape velocity, indicating the degree to which the core dominates these dynamics.

@@

The most noticeable distinction among models of different metallicities is the number of ejections, which decreases with increasing metallicity.
This is understandable, as in practice the lower metallicity models have greater BH populations than the higher metallicity models; quantitatively, the 1\% solar models have about $1.2 \times$ as many BHs at the time of first core collapse as the solar models, and maximum BH masses about $2 \times$ those of the same.
This weakens BSCO dynamics before and during first core collapse, as was first made visible in Figure \ref{fig:cmc_single_clusters}.
Because this effect takes place in the earlier stages of model evolution - namely, when mass segregation has had less time to separate lighter objects from the strong dynamics of heavier objects - this has the most significant effect of reducing the number of ejecta with $m < M_\odot$.
Notably, the different metallicity models have quite similar ejecta mass distributions after second core collapse, as the effect of metallicity on the masses of the remaining WDs and NSs is much less pronounced than on the now-ejected BHs.

@@

Having constrained the snapshots by $r_{\rm gc}$ and metallicity, we then choose one of an appropriate size and evolutionary state by comparing the snapshots with the model in normalized $\log M$ - $r_c / r_h$ space, where $M$ is the mass of the cluster in $M_\odot$, $r_c$ is the Spitzer core radius \citep{1987degc.book.....S}
\begin{equation}
    r_c = \sqrt{\frac{3 \sigma_c^2}{4 \pi \rho_c}},
\end{equation}
(where $\sigma_c$ is the central velocity dispersion and $\rho_c$ is the central density) and $r_h$ is the 3D half-mass radius of the snapshot (the latter two parameters are used because both are immediately accessible in the \citet{2018MNRAS.478.1520B} catalog and standard \CMC\ output).
The models and GCs are plotted in the non-normalized space of these parameters in the bottom plot of Figure \ref{fig:cmcs-mwgcs_scatter}.
The transformation to the normalized space is simply
\begin{equation}
    x_{\rm norm} = \frac{x - \bar{x}_{\rm \CMC}}{\sigma_{\rm x, \CMC}},
\end{equation}
where $\bar{x}_{\rm CMC}$ and $\sigma_{\rm x,CMC}$ are respectively the mean and standard deviation of the parameter $x$ over the complete set of \CMC\ snapshots.
The snapshot from the $r_{\rm gc}$ - [Fe/H] subset that is closest to the MWGC in the normalized $\log M$ - $r_c / r_h$ space is chosen to represent the GC; as an example, 47 Tuc ($r_{\rm gc} = 7.52$ kpc, $Z = 0.0019$; $\log M = 5.95$, $r_c / r_h,m = 0.125$ \citep{2018MNRAS.478.1520B, 2010arXiv1012.3224H}) is represented by a snapshot from model \texttt{N1.6e6\_rv2\_rg8\_Z0.002}, which at the 10.8 Gyr time of the snapshot has $\log M = 5.91$ and $r_c / r_h = 0.126$.

@@

Having chosen representative snapshots for all MWGCs, we now use our synthetic ejecta populations to compose a MW-like population of BSCO ejecta from GCs.
All orbit integrations in the galactic frame are done with \texttt{galpy}\footnote{http://github.com/jobovy/galpy} \citep{2015ApJS..216...29B}.
Setting the snapshot to the present-day galactic orbital parameters from \citet{2018MNRAS.478.1520B}, we back-integrate the orbit of each GC in the \texttt{galpy}\texttt{MWPotential2014} potential to the initialization time of the respective model.
We then locate each ejected object at the appropriate place in the GC orbit, using the time of BSCO encounter (naturally ignoring any ejections that occurred after the time of the chosen snapshot).
A random ejection direction is chosen in the GC rest frame for each object, and the velocities are then transformed to the galactic rest frame.
Finally, the orbits of all ejecta are then integrated to the present day in the same galactic potential.

@@

\begin{figure*}
    %\script{gc_orbit_ejections.py}
    \gridline{
        \fig{figures/gc_orbit_ejections_NGC_104.pdf}{0.49\textwidth}{}
        \fig{figures/gc_orbit_ejections_NGC_5139.pdf}{0.49\textwidth}{}
    }
    \gridline{
        \fig{figures/gc_orbit_ejections_NGC_6205.pdf}{0.49\textwidth}{}
        \fig{figures/gc_orbit_ejections_NGC_7089.pdf}{0.49\textwidth}{}
    }
    \caption{
        Plots showing the back-integrated orbits for some sample MWGCs (gray curves), and the points in the orbit where an object is ejected from the GC (scatter points).
        The color scale communicates the mass of the ejected star.
        The concentration of ejections in the first few orbits is clear, and an increased density of ejections when clusters are farther away from the galactic center is visible as well.
        Viewing the figure electronically makes it easier to find the few high-mass ejecta amid the abundance of lower mass objects.
    }
    \label{fig:gc_orbit_ejections}
\end{figure*}

@@

From a galactocentric perspective, the population of runaways is fairly isotropic in position, as can be seen in Figure \ref{fig:rgc-z}.
The distributions of distance from the galactic origin, and the local distributions of galactocentric velocity are similar when distributing over radius $r_{\rm gc}$ versus distance from the galactic plane $Z_{\rm gc}$.
Most runaways end up at a distance on the order of 10 kpc from the galactic center, and with a velocity on the order of 100-300~\kms in the galactocentric rest frame.
Few runaways make it past the 100 kpc mark, but those that do naturally retain the highest velocities, reaching upwards of 1000~\kms.

\begin{figure*}
    \script{rgc-z.py}
    \begin{centering}
        \includegraphics[width=0.9\linewidth]{figures/rgc-z_S.pdf}
        \caption{
            Histograms (top) and velocity quantiles (bottom) for the synthetic stellar ejecta.
            The left (right) plots show the profile over radial distance from the galactic center $r_{\rm gc}$ (distance from the galactic plane $Z$).
            The quantiles in the lower plots are calculated from the present-day velocities of our population.
            The red dots show the $r_{\rm gc}$ ($Z$) positions of the MWGCs from our composite \citet{2018MNRAS.478.1520B}+\citet{2010arXiv1012.3224H} catalog, arbitrarily spread in the vertical dimension.
        }
        \label{fig:rgc-z}
    \end{centering}
\end{figure*}

\begin{figure}
    \script{cmc_orbits_cdf.py}
    \begin{centering}
        \includegraphics[width=0.99\linewidth]{figures/cmc_orbits_cdf_S.pdf}
        \caption{
            Velocity distributions for our synthetic and some previous observational HVS catalogs.
            $v_{\rm rf}$ is the heliocentric radial velocity of the star as measured in the galactic rest frame.
        }
        \label{fig:cmc_orbits_cdf}
    \end{centering}
\end{figure}

\begin{figure}
    \script{mwej_rates.py}
    \begin{centering}
        \includegraphics[width=0.99\linewidth]{figures/mwej_rates.pdf}
        \caption{
            The evolution of the ejection rate for different components of our synthetic population.
            The blue (red) dashed line shows the estimated from observations for runaways from the galactic disc (HVSs from the galactic center), as reported by \citet{2015ARA&A..53...15B}.
        }
        \label{fig:mwej_rates}
    \end{centering}
\end{figure}

One result of note is that of the heliocentric radial velocity, $v_{\rm rf}$, distribution of the synthetic population.
\citet{2021arXiv211213864G} in part studied the galactic center origin of HVSs, and found that there is an apparent tension between the observed and predicted runaway populations; specifically, the predicted number of stars with velocities $\gtrsim 700~{\rm km~s^{-1}}$ was much higher than the observed rate from the HVS sample of \citet{2018ApJ...866...39B}.
The same work found that a decreasing star formation rate over last millions of years would reduce the recent (and thus more likely to be observed) HVS production rate enough to overcome the discrepancy.
GCs naturally follow this pattern (c.f. \S\ref{sec:gcdyns}), as the BSCO runaways studied here are produced at much higher rates when the models are young, and few HVSs are produced in the later stages of the cluster.
The resulting $v_{\rm rf}$ distribution is much closer to the observed distribution than that for the galactic center origin case, albeit skewed towards slightly smaller velocities (Figure \ref{fig:cmc_orbits_cdf}).
This latter difference grows when comparing to the HVS sample of \citet{2018ApJ...866..121H}, who focused on metal-poor stars, which are more akin to the objects that populate GCs.
